Artin Group Presentations Arising from Cluster Algebras

### 2014-07-20 Paper.tex
Intro and Background edits

In \cite{FZ02}, Fomin and Zelevinsky first introduced the concept of cluster algebras. This introductory paper focused on structural features of cluster algebras, specifically that, in a given cluster, when any cluster variable is viewed as a rational function in the variables of the given cluster, this cluster variable is a Laurent polynomial. \todo{The first sentence seems to be run on, and I'm very confused by it} Fomin and Zelevinsky defined cluster algebras in order to make further strides in the areas of representation theory, Lie theory, and total positivity. Since then, the study of cluster algebras has provided a motivation for applications in various other areas of mathematics, including quiver representations. Of particular interest were \textit{finite-type} cluster algebras, that is, cluster algebras whose variables are generated through mutation on a finite number of \textit{seeds}. In the sequel to their introductory paper (\cite{FZ03}), Fomin and Zelevinsky introduce the concept of \textit{mutation equivalence} between diagrams, proving that a connected graph is mutation equivalent to an oriented Dynkin diagram if and only if all mutation equivalent graphs have edge weights not exceeding 3. In particular, this proves that finite-type cluster algebras can be classified by Dynkin diagrams. \\
\indent Barot and Marsh extended Fomin and Zelevinsky's results in \cite{BM13}, providing a presentation of the reflection group associated to a Dynkin diagram with generators that correspond to elements of a companion basis associated to a seed of a finite-type cluster algebra. They also proved that this group presentation is invariant up to isomorphism under the mutation equivalence relation.

\begin{thm}\cite[Theorem 5.4]{BM13}
\begin{enumerate}
\item Let $\Gamma$ be a diagram of finite type and $\Gamma^{\prime} = \mu_k(\Gamma)$ the mutation of $\Gamma$ at vertex $k$. Then $W_{\Gamma} \cong W_{\Gamma^{\prime}}$.
\item Let $\mathscr{A}$ be a cluster algebra of finite type. Then the groups $W_{\Gamma}$ associated to the diagrams $\Gamma$ arising from the seeds of $\mathscr{A}$ are all isomorphic (to the reflection group associated to the Dynkin diagram associated to $\mathscr{A}$).
\end{enumerate}
\end{thm}

That is, given a diagram $\Gamma$ and a diagram mutation equivalent to $\Gamma$, denoted $\Gamma^{\prime} = \mu_k(\Gamma)$, they proved that $W_{\Gamma} \cong W_{\Gamma^{\prime}}$, where $W_{\Gamma}$ and $W_{\Gamma^{\prime}}$ are the group representations corresponding to $\Gamma$ and $\Gamma^{\prime}$, respectively. \\

@@

\section{Definitions and Notation}
\label{sec:main_definitions}

@@

Just as we can define mutations of the seeds of a cluster variable, we can also define mutations of a diagram associated to a cluster algebra of finite type by the following set of rules:
\begin{prop}\cite[Proposition 1.4]{BM13}
Let $B$ be a 2-finite skew-symmetrisable matrix. Then $\Gamma(\mu_k(B))$ is uniquely determined by $\Gamma(B)$ as follows: 
\begin{itemize}
\item Reverse the orientations of all edges in $\Gamma(B)$ incident with $k$ (leaving the weights unchanged)
\item For any path in $\Gamma(B)$ of form $i \xrightarrow{a} k \xrightarrow{b} j$ (i.e. with $a,b$ positive), let $c$ be the weight on the edge $j \rightarrow i$, taken to be zero if there is no such arrow. Let $c'$ be determined by $c'\geq 0$ and 
$c+c' =$ max$(a,b)$. 
Then $\Gamma(B)$ changes in a predetermined way, \todo{Say what this way is?} taking the case $c' = 0$ to mean no arrow between $i$ and $j$.
\end{itemize}
\end{prop}

@@

The rest of the paper will be devoted to building up analogous relations, defined in ~\ref{grp def} to prove a similar result in the case of Artin groups.

@@

\begin{rem}
Each Artin group has an associated Coxeter Group defined by adding in the additional relations $s_i^2 = 1$ for all $i.$ An Artin group is said to be of {\it finite type} if its associated Coxeter group is of finite type. To each Artin group of finite type we can assign it the same dynkin diagram which is assigned to the coxeter group associated to the Artin group.
\end{rem}

@@

For $\Gamma$ a diagram of finite type, we define the associated Artin Group as follows. The associated Artin group $W_\Gamma$ is generated by $s_i,$ where there is one $s_i$ for each vertex $i$ in $\Gamma.$ These generators are subject to the following relations
\begin{itemize}
	\item[(R2)] With $m_{ij}$ as defined in ~\ref{mij_defn}, for all $i \neq j,$ we add the relations
	$\langle s_i,s_j \rangle^{m_{ij}}= \langle s_j,s_i \rangle^{m_{ij}}.$

@@

\begin{rem}
In the above definition, the chordless cycle is ordered, which means that two cycles are considered different if they are cyclic rotations of eachother.
\end{rem}

\begin{rem}
We purposely include relations $(R2),(R3)$ but not $(R1)$ in order to make our relation labeling analagous to that of \cite{BM13} at the beginning of Section 3. Note that if we add the additional relation $(R1)$ as defined at the beginning of Section 3 of \cite{BM13}. Namely, if we add $s_i^2 = e$ for all vertices $i$ in $\Gamma,$ then we will precisely obtain the group $W_\Gamma$ as defined at the beginning of section 3 in \cite{BM13}.
\end{rem}

>>>>

In \cite{FZ02}, Fomin and Zelevinsky first introduced the concept of cluster algebras in order to make further strides in the areas of representation theory, Lie theory, and total positivity. Since then, the study of cluster algebras has provided a motivation for applications in various other areas of mathematics, including quiver representations. Of particular interest were \textit{finite-type} cluster algebras, that is, cluster algebras whose variables are generated through mutation on a finite number of \textit{seeds}. In the sequel to their introductory paper (\cite{FZ03}), Fomin and Zelevinsky introduce the concept of \textit{mutation equivalence} between diagrams, proving that a connected graph is mutation equivalent to an oriented Dynkin diagram if and only if all mutation equivalent graphs have edge weights not exceeding 3. In particular, this proves that finite-type cluster algebras can be classified by Dynkin diagrams. \\
\indent Barot and Marsh extended Fomin and Zelevinsky's results in \cite{BM13}, providing a presentation of the reflection group associated to a Dynkin diagram with generators that correspond to elements of a companion basis associated to a seed of a finite-type cluster algebra. They also proved that this group presentation is invariant up to isomorphism under the mutation equivalence relation.That is, given a diagram $\Gamma$ and a diagram mutation equivalent to $\Gamma$, denoted $\Gamma^{\prime} = \mu_k(\Gamma)$, they proved that $W_{\Gamma} \cong W_{\Gamma^{\prime}}$, where $W_{\Gamma}$ and $W_{\Gamma^{\prime}}$ are the group representations corresponding to $\Gamma$ and $\Gamma^{\prime}$, respectively.

%%\begin{thm}\cite[Theorem 5.4]{BM13}
%\begin{enumerate}
%\item Let $\Gamma$ be a diagram of finite type and $\Gamma^{\prime} = \mu_k(\Gamma)$ the mutation of $\Gamma$ at vertex $k$. Then $W_{\Gamma} \cong W_{\Gamma^{\prime}}$.
%\item Let $\mathscr{A}$ be a cluster algebra of finite type. Then the groups $W_{\Gamma}$ associated to the diagrams $\Gamma$ arising from the seeds of $\mathscr{A}$ are all isomorphic (to the reflection group associated to the Dynkin diagram associated to $\mathscr{A}$).
%\end{enumerate}
%\end{thm}

In our paper, we define $A_{\Gamma}$ to be the Artin group presentation arising from a cluster algebra, where $\Gamma$ is the diagram associated to the cluster algebra. We provide the necessary relations for the generators of the group, and show that these relations hold under mutations of chordless cycles in a diagram. Our main result is to show that this Artin group presentation is invariant up to isomorphism under the mutation equivalence relation. We state the result here, but present the detailed proof in Section \ref{sec:proof_of_main}.

\begin{thm}[Theorem \ref{thm:main}]
Let $\Gamma$ be a diagram of finite type, and let $\Gamma^{\prime} = \mu_k(\Gamma)$ be the mutation of $\Gamma$ at vertex $k$.  Then $A_{\Gamma} \cong A_{\Gamma^{\prime}}$.
\end{thm}

@@

\section{Background}
\label{sec:main_definitions}

@@

Just as we can define mutations of the seeds of a cluster variable, we can also define mutations of a diagram associated to a cluster algebra of finite type by the following set of rules:
\begin{prop}\cite[Proposition 1.4]{BM13}
Let $B$ be a 2-finite skew-symmetrisable matrix. Then $\Gamma(\mu_k(B))$ is uniquely determined by $\Gamma(B)$ as follows: 
\begin{itemize}
\item Reverse the orientations of all edges in $\Gamma(B)$ incident with $k$ (leaving the weights unchanged)
\item For any path in $\Gamma(B)$ of form $i \xrightarrow{a} k \xrightarrow{b} j$ (i.e. with $a,b$ positive), let $c$ be the weight on the edge $j \rightarrow i$, taken to be zero if there is no such arrow. Let $c'$ be determined by $c'\geq 0$ and 
$c+c' =$ max$(a,b)$. 
Then $\Gamma(B)$ changes as in Figure \ref{mutation}, taking the case $c' = 0$ to mean no arrow between $i$ and $j$.
\end{itemize}
\end{prop}

\begin{center}
\begin{tikzpicture}[baseline=-0.5ex]
\label{mutation}

	 \matrix[matrix of math nodes,column sep={25pt,between origins},row
    sep={40pt,between origins},nodes={asymmetrical rectangle}] at (0,0)
  {
    && |[name = u]|\bullet& \\
    %
    &|[name=dl]| \bullet & &|[name=dr]| \bullet \\
  };
\draw[->]
  (u) to node[right] {b} (dr)
  ;
  \draw[->]
  (dl) to node[left]{a}(u)
  ;
  \draw[->]
  (dr) to node[below]{c}(dl)
  ;
  \node [yshift = .4 cm] at (u){k};
    \node [yshift = -.4 cm,xshift = .4 cm] at (dr){j};
    \node [yshift = -.4 cm,xshift = -.4 cm] at (dl){i};
    \matrix[matrix of math nodes,column sep={25pt,between origins},row
    sep={40pt,between origins},nodes={asymmetrical rectangle}] at (4,0)
  {
    && |[name = u]|\bullet& \\
    %
    &|[name=dl]| \bullet & &|[name=dr]| \bullet \\
  };
\draw[->]
  (dr)to node[right] {b} (u) 
  ;
  \draw[->]
  (u) to node[left]{a}(dl)
  ;
  \draw[->]
  (dl) to node[below]{c'}(dr)
  ;
  \node [yshift = .4 cm] at (u){k};
    \node [yshift = -.4 cm,xshift = .4 cm] at (dr){j};
    \node [yshift = -.4 cm,xshift = -.4 cm] at (dl){i};

   \end{tikzpicture}
   \end{center}

@@

The rest of the paper will be devoted to building up analogous relations, defined in ~\ref{grp def} to prove a similar result in the case of Artin groups. For $\Gamma$ a diagram of finite type, we define the associated Artin Group as follows.

\begin{defn}

The associated Artin group $A_\Gamma$ is generated by $s_i,$ where there is one $s_i$ for each vertex $i$ in $\Gamma.$ These generators are subject to the following relations
\begin{itemize}
	\item[(R2)] With $m_{ij}$ as defined in ~\ref{mij_defn}, for all $i \neq j,$ we add the relations
	$\langle s_i,s_j \rangle^{m_{ij}}= \langle s_j,s_i \rangle^{m_{ij}}.$

 We will use these group relations to prove the mutation invariance of $A_{\Gamma}$ in Section \ref{sec:proof_of_main}

@@

\begin{rem}
Each Artin group has an associated Coxeter Group defined by adding in the additional relations $s_i^2 = 1$ for all $i.$ An Artin group is said to be of {\it finite type} if its associated Coxeter group is of finite type. To each Artin group of finite type we can assign it the same Dynkin diagram which is assigned to the Coxeter group associated to the Artin group.
\end{rem}

@@

Recall from Section \ref{sec:main_definitions} that for $\Gamma$ a diagram of finite type, we define the associated Artin Group as follows. The associated Artin group $A_\Gamma$ is generated by $s_i,$ where there is one $s_i$ for each vertex $i$ in $\Gamma.$ These generators are subject to the following relations
\begin{itemize}
	\item[(R2)] With $m_{ij}$ as defined in ~\ref{mij_defn}, for all $i \neq j,$ we add the relations
	$\langle s_i,s_j \rangle^{m_{ij}}= \langle s_j,s_i \rangle^{m_{ij}}.$

@@

\begin{rem}
In the above definition, the chordless cycle is ordered, which means that two cycles are considered different if they are cyclic rotations of each other.
\end{rem}

\begin{rem}
We purposely include relations $(R2),(R3)$ but not $(R1)$ in order to make our relation labeling analogous to that of \cite{BM13} at the beginning of Section 3. Note that if we add the additional relation $(R1)$ as defined at the beginning of Section 3 of \cite{BM13} (namely, if we add $s_i^2 = e$ for all vertices $i$ in $\Gamma$), then we will precisely obtain the group $W_\Gamma$ as defined at the beginning of section 3 in \cite{BM13}.
\end{rem}
